MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-22 chapters/MetaTracts.tex
Updated text

\section{Introduction and Motivation}
\label{sec:intro}
% Modern industry is constantly driving their research towards novel lightweight components. For these high-tech applications, the industry unites behind the mantra of function orientation, high integration as well as cost and eco-efficiency. A promising type of material, showing high specific stiffness, high specific strength, and high corrosion resistance, allows to integrate today's demands of industry. Advanced composites and in particular carbon fiber reinforced polymers (CFRP) were successfully introduced in aeronautic and automotive applications and within the recent years their share is rapidly rising. To increase the usage, not only structural but also complex primary structures and  highly loaded components have been manufactured from CFRP. As for today,  besides commercial aerospace and automotive industry, CFRP is used in wide range of industries, e.g., space/satellite, marine, sporting goods, automotive, civil engineering or wind energy.

%REPLACED BY 
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components. The material of choice of a strongly growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing more and more parts made of conventional materials such as aluminum or steel. But besides the aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

In general, CFRP materials show highly desired characteristics as high specific stiffness, high specific strength, high corrosion resistance at considerably lower weight compared to conventional materials. At the same time highly complex and integrated components may be produced from CFRPs, which were previously impossible to manufacture. For example even primary structures and highly loaded components in aeronautics are been increasingly manufactured from CFRP.

Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components:

The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics. To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used. In special applications additional filler materials, such as colors or other particles, may be integrated. 
%The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.

The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is yet the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On the one hand ultrasonic testing provides a quick and cost-efficient overview. But on the other hand it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which taken throughout a full rotation of the specimen. In industrial XCT, the typical scanning geometry foresees the specimen to be placed on the rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays get attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image), depending on its principle e.g. by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is finally reconstructed to a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below $500 nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of the individual fibers.
%of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-Ray computed tomography (XCT) with regard to larger detectors, larger field of views and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.

@@

Extracting the answer to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts, ``geometric structure" and ``spatial context". Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.

The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data. Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provides details of our approach. Section \ref{sec:results} details experiments and Section~\ref{sec:param_choices} explains our  parameter choices. We end with conclusions.

@@

Diffusion Magnetic Resonance Imagining (dMRI, also referred to as Diffusion Tensor Imaging (DTI)) has gained popularity in medical diagnosis within recent years. Its main clinical application is found in the study and treatment of neurological disorders. DTI may reveal abnormalities in white matter fiber structure and allows for visualizing the organization of white fibers in the human brain as well as the brain connectivity.
A variety of algorithms have been proposed aiming to generate fiber-tract trajectories. In general, these reconstructions of fiber trajectories are then clustered in to bundles which are expected to be related anatomically or spatially. 
We broadly divide the related work on DTI into two parts which are of immediate relevance to our proposed solution: fiber tracking and fiber clustering. Furthermore methods analyzing the second order local structure are also discussed. Finally, we review the current state of the art in the visual analysis of fiber reinforced composites.

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
In detecting fiber tracts as shown by Basser et al.~\cite{Basser2000},~\cite{Basser2002} and Mori et al.~\cite{Mori2002},~\cite{Mori1999}, the key idea is based on the assumption that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel. In these techniques, continuous tracts are created by propagating a virtual particle from a starting point until they reach a termination criterion. This is usually done by solving a second-order Runge-Kutta integration. The assumption is that the principal diffusion direction at each discrete location can be interpolated to from a continuous velocity field. One of the short comings of the process is that because of the local decision-making process, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2000},~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods. While, energy minimization approaches find the most favorable trajectory connecting two given endpoints.

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
%!!!!! TODO REFINE FIBER CLUSTERING !!!!!
%The next step is to identify  anatomically relevant fiber bundles. This is achieved through fiber clustering.
The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles. Extensive research has been done on automatic DTI fiber clustering methods
% (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).

The clustering approach divides the problem into two parts. The first part deals with finding a suitable proximity measure and the second deals with the choice of the clustering method. Pairwise proximity measures include, endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of the above distance so that fibers that are close for certain distance and then diverge are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering (see ~\cite{Moberts2005},~\cite{Zhang2008}) and those using spectral clustering (see~\cite{jonasson2005},~\cite{ODonnell2007},~\cite{Brun2004})

Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps proposed by Belkin and Niyogi~\cite{Belkin01} can be applied to the problem of organizing fiber trace data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber traces. In the ideal case fiber traces which belong to the same bundle must remain ``close" to each other in the lower dimensional space.Westin et al.~\cite{westinMEDIA02} also uses spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} runs k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}. 
These approaches build on the assumption that proximity measure that compare DTI fiber trajectories can also represent anatomical relationship. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are joined. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

The difficulties in hierarchical clustering include: first, computing all pairs distances for tracts to generate the distance matrix (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times). Second, choosing the ``correct'' distance measure to compare tracts. Thirdly, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

\subsection{Second order local structure}
%!!!!! TODO REFINE SECTION !!!!!
Unlike DTI, we do not have diffusion tensor data. Instead, we have a scalar volume with tubular structures embedded in them. Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement as proposed by Frangi et al.~\cite{Frangi1998} and Sato et al.~\cite{Sato1997}. Frangi et al.~\cite{Frangi1998} introduced a method based on studying the eigenvalues of the local second order structure namely the Hessian matrix, specifically for the purposes of developing vessel enhancement filters.

@@

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the datatype is uint8. It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis. Figure~\ref{fig:data-char}(D) shows the zoomed in version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
\includegraphics[width=0.5\textwidth]{imagesMT2014/image3b.png}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps. MetaTracts generations Sec.~\ref{sec:approach} and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.

@@

The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] which represents a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel. The principal directions, in which the local second order structure of the image can be decomposed are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.

@@

\begin{enumerate}
\item{Each MetaTract is associated with a continuous set of cylinders.}
\item{Each MetaTract is associated with a start point which is a grid vertex.}
\item{Each cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
\item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
\item{Each cylinder is locally parallel to the local orientation at it's start point.}
\end{enumerate}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length)}
	\label{fig:length_distribution}
\end{figure}

>>>>

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The material of choice of a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing  parts made of conventional materials such as aluminum or steel. CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength, high corrosion resistance. CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
At the same time highly complex and integrated components which were previously  impossible to manufacture may be produced from CFRP's.
For example even primary structures and highly loaded components in aeronautics are being increasingly manufactured from CFRP.

Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components.
A matrix on the one side, which acts as a bonding component and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. 
The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On the one hand ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly being applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. In industrial XCT, the typical scanning geometry places the specimen on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays get attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below $500 nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work, we focus on datasets with larger viewports and lower resolutions where the single carbon fibers (filaments) are not or are  barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of  individual fibers.
%of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-Ray computed tomography (XCT) with regard to larger detectors, larger field of views and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.

@@

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts, ``geometric structure" and ``spatial context". Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.

The paper is structured as follows: Section~\ref{sec:prev_work} reviews the related work,  Section \ref{sec:char_data} describes the   characteristics of our data. Sections \ref{sec:approach},~\ref{subsec:fiber-bundles},~\ref{sec:vis} provides details of our approach. Section \ref{sec:results} details experiments, section~\ref{sec:user_eval} summarizes our user evaluation results. Section~\ref{sec:param_choices} explains our  parameter choices. We end with conclusions in section~\ref{sec:conclusions}.

@@

Diffusion Magnetic Resonance Imagining (dMRI, also referred to as Diffusion Tensor Imaging (DTI)) has gained popularity in medical diagnosis within recent years. Its main clinical application is found in the study and treatment of neurological disorders. DTI may reveal abnormalities in white matter fiber structure and is used for visualizing the organization of white fibers in the human brain and  brain connectivity.
A variety of algorithms have been proposed aiming for generating fiber-tract trajectories. In general, these reconstructions of fiber trajectories are  clustered in bundles which are expected to be related anatomically or spatially. 
We broadly divide the related work on DTI into two parts that are of immediate relevance to our proposed solution: fiber tracking and fiber clustering. We discuss  methods for analyzing the second order local structure. Finally, we review the current state of the art in the visual analysis of fiber reinforced composites.

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}. Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion. 
 This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to from a continuous velocity field. 
 Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2000},~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori1999},~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Corouge2004, Brun2003, Brun2004, Zhang2008, westinMEDIA02}. 
Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering ~\cite{Moberts2005, Zhang2008} and those using spectral clustering ~\cite{jonasson2005, ODonnell2007, Brun2004}.

Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}) can be applied to the problem of organizing fiber trace data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber traces. In the ideal case fiber traces that belong to the same bundle must remain ``close" to each other in the lower dimensional space. Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}. 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationship. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are joined. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

There are a number of difficulties in hierarchical clustering. First, computing all pairs distances for tracts to generate the distance matrix is time consuming (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times). Second, a ``correct'' distance measure to compare tracts must be chosen. Thirdly, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

\subsection{Second order local structure}
%!!!!! TODO REFINE SECTION !!!!!
Unlike DTI, we do not have diffusion tensor data. Instead, we have a scalar volume with tubular structures embedded in them. Analyzing curvilinear structures in volumetric images has been utilized for a variety of purposes including center line extraction~\cite{Bouix2005} and vascular image enhancement as proposed by Frangi et al.~\cite{Frangi1998} and Sato et al.~\cite{Sato1997}. Frangi et al.~\cite{Frangi1998} introduced a method based on studying the eigenvalues of the  Hessian matrix specifically for the purposes of developing vessel enhancement filters.

@@

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the data type is unsigned 8 bit integer. Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis. Figure~\ref{fig:data-char}(D) shows the magnified version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\begin{figure}[tb]
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps. MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.

@@

The input to the first stage of our pipelines, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location becomes associated with a vector representing the local orientation and a real value [0,1] which represents a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel. The principal directions, in which the local second order structure of the image can be decomposed, are given by the eigen decomposition of the Hessian matrix which describes the local curvature.
The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.

@@

\begin{enumerate}
\item{Each MetaTract is associated with a continuous set of cylinders.}
\item{Each MetaTract is associated with a start point which is a grid vertex.}
\item{Each individual cylinder in a MetaTract has a constant length, a constant radius and a start point (which is also a grid vertex).}
\item{Each cylinder in a MetaTract (except for the first one) is connected to the previous cylinder at its start point.}
\item{Each cylinder is locally parallel to the local orientation at it's start point.}
\end{enumerate}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

\section {Fiber Bundle Generation}
\label{subsec:fiber-bundles}
